---
title: "Regression"
author: "Rahul Singh"
date: "12/29/2017"
output: html_document
---
---
title: "ML Assignment Linear Regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
getwd()
# setwd("/users/anchaljaiswal/Downloads")
library(dplyr)
library(ggplot2)
library(car)
library(gridExtra)
library(tidyr)
library(GGally)
getwd()
inp=read.csv("Houses_for_rent_madrid_v5.csv")
df=data.frame(inp)
str(df)
```


## R Markdown

```{r}
mean_price=mean(df$Rent)
df$Outer=factor(df$Outer)
df$Elevator=factor(df$Elevator)
df$Penthouse=factor(df$Penthouse)
df$Cottage=factor(df$Cottage)
df$Duplex=factor(df$Duplex)
df$Semi.detached=factor(df$Semi.detached)

df = df %>%
  mutate(floor_cat=ifelse(Floor<=5,1,ifelse(Floor>=6 & Floor<=10,2,3))) %>%
  mutate(bed_type = ifelse(Bedrooms<=2,1,ifelse(Bedrooms>=3 & Bedrooms<=4,2,3)))
df$floor_cat=factor(df$floor_cat)
df$Area=factor(df$Area)
df$bed_type=factor(df$bed_type)
str(df)

```

```{r}
head(inp)
```

# variable analysis

```{r}
# number of listings by area
ggplot(df,aes(x=reorder(Area,Area,
                     function(x)-length(x)), fill = Region)) +
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("Areas")

table(df$Area)
```


```{r}
# how many zones does each area have?
df %>%
  group_by(Area) %>%
  summarise(unique_zones = length(unique(Zone))) %>%
  arrange(desc(unique_zones)) %>%
  ggplot(aes(x = Area, y = unique_zones)) +
  geom_col() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
# rent price distribution by area?
ggplot(df, aes(x = Area, y = Rent, fill = Region)) +
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = mean(df$Rent), color = I('red'), linetype = 2)
```

```{r}
# rent prices by region?
ggplot(df, aes(x = Region, y = Rent, fill = Region)) +
  geom_boxplot() +
  geom_hline(yintercept = mean(df$Rent), color = I('red'), linetype = 2)
```

```{r}
# rent prices for number of bedrooms?
grid.arrange(
  ggplot(df, aes(x = factor(Bedrooms), y = Rent)) +
    geom_boxplot() +
    ggtitle("Average price by bedroom number"),

  ggplot(df,aes(x = Bedrooms)) +
    geom_bar() +
    ggtitle("Number of properties by bedroom number") +
    scale_x_continuous(breaks = seq(0,8,1)),
  ncol = 2
)
```
Price increases in general with more bedrooms, but with larger variance too. One explanation would be that sometimes there are cheap student apartments with large number of bedrooms available, and thus while the property is large, it can hardly be considered a 'valuable' one. 

```{r}
# number of bedrooms by area
df %>%
  group_by(Area) %>%
  ggplot(aes(x = Area, y = Bedrooms, fill = Region)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = mean(df$Bedrooms))
```



```{r}
# type of house price difference?
grid.arrange(
ggplot(inp, aes(x = factor(Studio), y = Rent)) +
  geom_boxplot() +
  ggtitle('Rent: studio'),

ggplot(inp, aes(x = factor(Penthouse), y = Rent)) +
  geom_boxplot() +
  ggtitle('Rent: penthouse'),
ncol = 2
)
```


```{r}
grid.arrange(
ggplot(inp, aes(x = factor(Cottage), y = Rent)) +
  geom_boxplot() +
  ggtitle('Rent: cottage'),

ggplot(inp, aes(x = factor(Duplex), y = Rent)) +
  geom_boxplot() +
  ggtitle('Rent: duplex'),

ggplot(inp, aes(x = factor(Semi.detached), y = Rent)) +
  geom_boxplot() +
  ggtitle('Rent: semi detached'),

ncol = 3
)
```



```{r}
ggplot(inp, aes(x = Inside_M30)) +
  geom_bar()

mean_region_m30 <- inp %>%
  group_by(Inside_M30, Region) %>%
  summarise(mean_rent = mean(Rent))

grid.arrange(
ggplot(subset(mean_region_m30, Inside_M30 == 0), aes(x = Region, y = mean_rent, fill = Region)) +
  geom_col() +
  ggtitle("Mean rent regions out of M30"),

ggplot(subset(mean_region_m30, Inside_M30 == 1), aes(x = Region, y = mean_rent, fill = Region)) +
  geom_col() +
  ggtitle("Mean rent regions inside M30"),

ncol = 2
)
```


```{r}
grid.arrange(
  ggplot(subset(inp, Inside_M30 == 0),
         aes(x = Region, y = Rent)) +
    geom_boxplot() +
    stat_summary(fun.y = mean, geom = 'point', color = I('red')) +
    ggtitle("Rent among Regions out of M30") +
    scale_y_continuous(breaks = seq(0, 17500, 2500), limit = c(0, 17000)),
  
  ggplot(subset(inp, Inside_M30 == 1),
         aes(x = Region, y = Rent)) +
    geom_boxplot() +
    stat_summary(fun.y = mean, geom = 'point', color = I('red')) +
    ggtitle("Rent among Regions inside M30")+
    scale_y_continuous(breaks = seq(0, 17500, 2500), limit = c(0, 17000)),
  
  ncol = 2
)
```
While the North and South show great variation in rent depending on whether the area is in or out of M30, The South and West, while the rent prices are still higher for areas inside M30, arguably have 'less' variation in Rent prices, whether it is in or out of the M30. In fact, West seems to have some few outliers with higher Rent prices for areas out of M30, indiciating that there may be a affluent suburb in this region out of the busy areas.

## Linear Regression

```{r pressure, echo=FALSE}
fit2= lm(data=df, Rent ~ Sq..Mt. + floor_cat + Penthouse + Cottage  + Region + Outer)
summary(fit2)
```


```{r}
fit3= lm(data=df, Rent ~ Sq..Mt. + floor_cat + Cottage + Region + Duplex + Inside_M30)
summary(fit3)
```


#check correlation between IVs and DV

```{r}

df1 = subset(df, select = -c(Address))
df1 = subset (df, select = -c(Address, Zone, Area, Id))
ggpairs(df1)

```

## VERIFY STATISTICAL ASSUMPTIONS

```{r}

#heterocedasticity

qplot(predict(fit3), rstandard(fit3)) + geom_hline(yintercept = 0) + geom_hline(yintercept = 2, colour=I("red"), alpha=I(0.05)) + geom_hline(yintercept = -2, colour=I("red"), alpha=I(0.05))


qplot(predict(fit3), rstandard(fit3)) + geom_hline(yintercept = 0) + geom_hline(yintercept = 2, colour=I("red"), alpha=I(0.05)) + geom_hline(yintercept = -2, colour=I("red"), alpha=I(0.05))

#normality of residuals

q1 = qplot(rstandard(fit2), geom="blank") + geom_histogram(aes(y=..density..), colour=I("gray"), binwidth=0.5)+ stat_function(fun=dnorm, args=list(mean=0, sd=1), colour=I("red"), alpha=I(0.5))

q2 = qplot(sample=rstandard(fit2)) +
  geom_abline(slope=1,intercept=0)

grid.arrange(q1, q2, nrow=1)



q13 = qplot(rstandard(fit3), geom="blank") + geom_histogram(aes(y=..density..), colour=I("gray"), binwidth=0.5)+ stat_function(fun=dnorm, args=list(mean=0, sd=1), colour=I("red"), alpha=I(0.5))

q23 = qplot(sample=rstandard(fit3)) +
  geom_abline(slope=1,intercept=0)

grid.arrange(q13, q23, nrow=1)


#independence of residuals

durbinWatsonTest(fit2)
durbinWatsonTest(fit3)


#Global Linear Model Assumptions

gvlma(fit2)
gvlma(fit3)


```

So, apparently our model is heterocedastic but residuals are not independent and they don't follow a normal distribution.




## REGRESSION DIAGNOSTIC

```{r}
#Multicollinearity

vif(fit2)
vif(fit3)

#Outliers

outlierTest(fit2)
outlierTest(fit3)

#Influenters

influencePlot(fit2, id.method = "identify")
dfbetaPlots(fit2, id.method = "identify")


```

There's not collinearity.
There are outliers that should be taken into account.



